%
\hsection{Join-based Select and Views}%
%
So far, things are going well.
On the positive side, we have learned that we can divide the data of factory into separate aspects and then store it into different tables.
This is nice.
We also learned that we can ensure the integrity of the data inside a single table in various ways.
For example, we can use datatypes that strictly enforce their domains.
\sqlilIdx{DECIMAL}, for example, makes sure that numbers are represented exactly without loss of precision to a fixed number of decimals~\cite{PGDG:PD:NT}.
\sqlilIdx{DATE} ensures that valid dates based on the Gregorian calendar are entered~\cite{PGDG:PD:HU,G1582IG}.
We can also define constraints, ranging from ensuring that all columns are entered~(\sqlilIdx{NOT NULL}) over preventing name clashes~(\sqlilIdx{UNIQUE}) to sanity checks~(via \sqlil{CHECK} constraints\sqlIdx{CONSTRAINT!CHECK})~\cite{PGDG:PD:C}.
We also learned that we can even ensure the consistency of the data in our \db\ across different tables via the \sqlilIdx{REFERENCES} constraint~\cite{PGDG:PD:FK}.
So we cannot just have tables, they can have strongly-typed columns and the data integrity can be enforced throughout the complete \db.
This puts \pglspl{rdb} well ahead the spread sheets produced by the likes of \microsoftExcel\ or \libreofficeCalc.

However, on the negative side, we have to admit that using \pglspl{db} is more complicated.
And we are certainly losing in terms of readability of our data.
Indeed, the meaning of the \sqlil{demand} data shown in \cref{exec:factory:insert_into_table_demand} is not obvious without knowing the contents of the other tables.
Now we will see that we can also \emph{use} these contents of other tables directly in our queries, to produce clear and human-readable output.%
%
\hsection{Joining Tables}%
%
We want to have a list of all of our customers, and for each customer we want to see the \sqlil{id} values of all of the demands~(orders) they have issued.
We build the query used in \cref{lst:factory:select_customer_demand} step-by-step.
The result of our query should have two columns.
The first column should hold the customer names.
Let's call it \sqlil{name}.
The second column should hold the \sqlil{id} of the order the customer made.
Let's call this column \sqlil{demand_id}.
Now if a customer issued multiple orders, then they should appear multiple times in this table, once for each demand.
If a customer did not issue any orders, then they should still appear, but only once and with a \sqlilIdx{NULL} value in the \sqlil{demand_id} column.

This means that first, we need all the customer names.
A \sqlil{SELECT name FROM customer;} would do this for us.
Now we need to cross-reference the customer \sqlil{id} values in table \sqlil{customer} with the column \sqlil{customer} in table \sqlil{demand}.
We also want to list the customers who did not issue any order.%
%
\hsection{\texttt{LEFT JOIN}}%
%
\begin{sloppypar}%
What we want to do is to apply a so-called \sqlil{LEFT JOIN}\sqlIdx{JOIN!LEFT}, also called \sqlIdx{LEFT OUTER JOIN}\sqlIdx{JOIN!LEFT OUTER}, of the table \sqlil{customer} to the table \sqlil{demand}~\cite{PGDG:PD:JT}.
The syntax of a \sqlil{LEFT JOIN}\sqlIdx{JOIN!LEFT} is as follows:%
\end{sloppypar}%
%
\sqlSyntax{syntax/left_join.sql}%
%
\gitLoadAndExecSQL{factory:select_customer_demand}{}{factory}{select_customer_demand.sql}{factory}{boss}{superboss123}%
\listingSQLandOutput{factory:select_customer_demand}{%
Get the per-customer demands.%
}{}%
%
In this query, usually \sqlil{Table_1} has primary key~\sqlil{a}, which is referenced as foreign key from column~\sqlil{b} of \sqlil{Table_2}.
In our case, \sqlil{Table_1} would be the table~\sqlil{customer}.
Instead of~\sqlil{a}, the primary key column is~\sqlil{id}.
\sqlil{Table_2} corresponds to table~\sqlil{demand}.
Column~\sqlil{customer} in table~\sqlil{demand} then corresponds to~\sqlil{Table_2.b}.
Notice that we can prefix column names with table names to add clarity:
\sqlil{Table_1.a} is the same as column~\sqlil{a} of table~\sqlil{Table_1}.

The important part of the \sqlil{LEFT JOIN}\sqlIdx{JOIN!LEFT} is the \sqlilIdx{ON} condition.
It can be an arbitrarily complex condition, involving \sqlilIdx{AND} and \sqlil{OR} and whatnot.
However, in the simplest form, it just picks a column from \sqlil{Table_1}, here \sqlil{Table_1.a} and says that its value must be the same as a column in \sqlil{Table_2}, here \sqlil{Table_2.b}.%
This condition then applies to the rows that are returned, with the only exception that, if there is no entry in \sqlil{Table_2} for some value of \sqlil{Table_1.a}, then \sqlil{NULL} will be returned for all fields used from \sqlil{Table_2} in the corresponding result row.

We apply this to our situation.
As said, \sqlil{Table_1} is obviously \sqlil{customer} and \sqlil{Table_2} is \sqlil{demand}.
The thing that we want to select is \sqlil{customer.name} and the corresponding values of \sqlil{demand.id}.
The fields that need to match are \sqlil{customer.id} and \sqlil{demand.customer}.%
%
\begin{sloppypar}%
In \cref{lst:factory:select_customer_demand}, we \sqlil{SELECT customer.name, demand.id} and we do so \sqlil{FROM customer LEFT JOIN demand ON (customer.id = demand.customer)}.
When executed, this query goes through the table~\sqlil{customer}.
For each record, it will take the field~\sqlil{id}, i.e., \sqlil{customer.id}.
It will then search for any row in the table~\sqlil{demand} whose field \sqlil{customer}, i.e., \sqlil{demand.customer} has the same value.
For any such row, it will write a row to the output with the \sqlil{name} of the customer and the \sqlil{id} of the row in \sqlil{demand}.
If and only if no such row exists for a customer, i.e., if the customer did not yet make any purchase, it writes a row to the output with the \sqlil{name} of the customer and \sqlilIdx{NULL} as the \sqlil{demand}~id.
This means that we get the demands associated with each customer name.
We also see which customer did not make any purchase.%
\end{sloppypar}%
%
To clean up the output, we rename \sqlil{customer.name AS name}\sqlIdx{AS} and \sqlil{demand.id AS demand_id}\sqlIdx{AS} to make it clearer which name and ID~values will be displayed.
We also add an \sqlil{ORDER BY name}\sqlIdx{ORDER BY} to sort the output by customer name.

The result of this query, shown in \cref{exec:factory:select_customer_demand}, has nine rows.
We have eight demands, which are associated with customers Bebba, Bebbo, and Bibbo.
Bobbo did not yet make any purchase, so he appears as a single row with \sqlilIdx{NULL} associated as \sqlil{demand_id}.
The \psql\ \pgls{client} just leaves \sqlil{demand_id} blank in the output.

We now have cross-referenced two tables!
Let us take this a step farther and ask \inQuotes{How many orders did each customer make so far?}
We again answer this question in \cref{lst:factory:select_customer_demand}.
Doing this is rather easy:
First, we replace \sqlil{demand.id AS demand_id} in the query with \sqlil{COUNT(*) as demands}\sqlIdx{COUNT}.
We already learned that \sqlilIdx{COUNT} just counts rows over groups.
But in order to let it count meaningfully, we need to create these groups.
All we have to do for this is to add a \sqlil{GROUP BY name}\sqlIdx{GROUP BY} to the query.
All rows that share a \sqlil{name} go into the same group.
\sqlilIdx{COUNT} now counts the rows in these groups.
The output in \cref{exec:factory:select_customer_demand} shows us that Bebba made two orders, Bebbo~four, and Bibbo also two.
Bobbo so far has zero orders in his name.

With this, we now know how to cross-reference tables.
We have essentially solved the basic annoyance mentioned before:
Yes, our data is distributed over different tables and without knowing what product a product~id refers to and which customer corresponds to which customer~id, it is hard to understand anything.
However, we can just join tables pull all the information together in a query.%
%
\FloatBarrier%
\endhsection%
%
\hsection{\texttt{INNER JOIN}}%
%
We are now able to cross-reference different tables and to pull data together.
Let us create a query where, for each demand record, we see the customer name, the product name, the product price, the ordered amount, and the order date.
The result of this query would then be a clear and human readable overview of the business transactions of our company.
Based on the previous example, where we joined the table \sqlil{customer} with the table \sqlil{demand}, we anticipate that we will need two \inQuotes{joins} now.
We will need to combine the \sqlil{demand} table with the \sqlil{customer} table and with the \sqlil{product} table.
The query will therefore probably be more complicated.

Let us construct the new query.
We first define what we want in a way more close to \sql:
\emph{\inQuotes{For each row in table~\sqlil{demand}, we need the corresponding row in table~\sqlil{customer} and the corresponding row in table~\sqlil{product}.}}
We know that there always exists exactly one \sqlil{customer} row, because we defined the foreign key \sqlilIdx{REFERENCES} constraint that enforces this.
We also know that there always exists exactly one \sqlil{product} row, because we defined the foreign key \sqlilIdx{REFERENCES} constraint that enforces this.
So it cannot be that we have a row \sqlil{demand} but no corresponding row in \sqlil{customer} or \sqlil{product}.
We already know that it is totally possible the other way around, i.e., that there is a row in \sqlil{customer} without any related row in \sqlil{demand} -- but this is not important here.

Given this information, we \emph{could} use the \sqlil{LEFT JOIN}\sqlIdx{JOIN!LEFT} from before.
It would generate rows also for rows of \sqlil{demand} that do not match to any \sqlil{customer} or \sqlil{product}.
But such rows do not exist anyway.
However, what we actually would like to use here is an \sqlil{INNER JOIN}\sqlIdx{JOIN!INNER}~\cite{PGDG:PD:JT}.
The \sqlil{INNER JOIN} only creates output rows when the rows of all involved tables match each other.
While this is not an issue here, it is an important aspect in many other situations.%
%
\sqlSyntax{syntax/inner_join.sql}%
%
But how do we go about joining \emph{three} tables?
We can chain arbitrary join expressions.
Thus, to merge data from three tables, we simply chaining two \sqlil{INNER JOIN} expressions:%
%
\sqlSyntax{syntax/inner_join_2.sql}%
%
\gitLoadAndExecSQL{factory:select_sale_data}{}{factory}{select_sale_data.sql}{factory}{boss}{superboss123}%
\listingSQLandOutput{factory:select_sale_data}{%
Selecting the sales data by joining the three tables~\sqlil{demand}, \sqlil{customer}, and \sqlil{product}.%
}{}%
%
\begin{sloppypar}%
Equipped with the understanding of this concept, we begin by doing basically the same thing as before:
We \sqlil{SELECT <something> FROM demand} and then \sqlil{INNER JOIN customer ON (customer.id = demand.customer)}.
For each row in \sqlil{demand}, this gives us the corresponding row in \sqlil{customer}.
It also would drop each row in \sqlil{demand} for which no row in \sqlil{customer} exists from the output, but that cannot happen anyway.%
\end{sloppypar}%
%
\begin{sloppypar}%
Now we simply \emph{chain} the next \sqlil{INNER JOIN} by writing it directly after that!
We write \sqlil{INNER JOIN product ON (product.id = demand.product)}.
This will also select the corresponding row from table \sqlil{product}.
If such a row does not exist, then the whole current combined row from \sqlil{demand} and \sqlil{product} would be dropped from the output.
Thanks to our \sqlilIdx{REFERENCES} constraints in table \sqlil{demand}, this, again, can never happen anyway.%
\end{sloppypar}%
%
We have nicely pulled the customer and product data for each demand order.
Having all the data at hand, we can choose the columns that we want to output and replace the \sqlil{<something>} we first nonchalantly wrote when beginning to design the query.
We want to get the customer name and product name.
Notice how both the table \sqlil{customer} and the table \sqlil{product} have a column called~\sqlil{name}.
To be able to tell them apart, we will not write \sqlil{name} for the customer name like in the previous big query that we designed.
Instead, we will select the customer name as \sqlil{customer.name AS customer_name}.
For the product name, we write \sqlil{product.name AS product_name}.
Thus, there can never be any confusion here, because we refer to these columns as \sqlil{customer.name} and \sqlil{product.name}, respectively.
Next we want the list product price and amount associated with an order, i.e., \sqlil{product.price AS price} and \sqlil{demand.amount AS amount}.
Finally, we also return the order date and therefore add column \sqlil{demand.ordered as ordered}.

For good measures, we add an \sqlilIdx{ORDER BY} clause and sort the output by \sqlil{customer_name}.
If two entries have the same \sqlil{customer_name}, then we sort the one with the earlier order date (\sqlil{ordered}) first.
If two entries have the same \sqlil{customer_name} \emph{and} \sqlil{ordered} date, then we sort the one with the lexicographically smaller \sqlil{product_name} first.
We try to resolve further draws by \sqlil{price} and \sqlil{amount}.

The full query is shown in \cref{lst:factory:select_sale_data}.
The output in \cref{exec:factory:select_sale_data} shows the eight orders issued.
However, unlike the raw data from table~\sqlil{demand}, the output is now clear and easy to read.%
%
\FloatBarrier%
\endhsection%
%
\endhsection%
%
\hsection{Views as Virtual Tables}%
%
\gitLoadAndExecSQL{factory:create_view_sale}{}{factory}{create_view_sale.sql}{factory}{boss}{superboss123}%
\listingSQLandOutput{factory:create_view_sale}{%
Creating a view, i.e., a stored \sql\ query that can be treated like a table, to list the demands in human-readable form.%
}{}%
%
The very clear sales information that we just saw is very useful for our factory.
Indeed, it is likely that we will need this query structure several times.
For example, we probably want to be able to query all the orders that were made in a specific year.
This would basically extend the same query with a \sqlilIdx{WHERE} clause attached that limits the demand dates.
Or maybe we want to sum up the cash flow per customer, to see who our most lucrative customers are and to maybe make them some special offers
Or maybe we want to summarize the total sales per product.
This would tell us which of our products sell well and which don't.

Every time we would re-write the query, just a bit differently.
Of course, this is not a good approach.
On the one hand, every time we write a query, there is a chance that we introduce a mistake.
On the other hand, if we ever decided to change the structure of \db, maybe by allowing multiple items per demand, then we are stuck with a heap of queries that all need to be changed.

Luckily, the \sql\ language and \pglspl{rdb} offer us a tool for this situation:
We can store the query as so-valled \emph{view}~\cite{PGDG:PD:CV}.
A view is basically a stored query that we can work with exactly as if it was a table.
We can apply other \sqlil{SELECT}\sqlIdx{SELECT{\idxdots}FROM} queries on top of it.
Thus, all we have to do is to store this query as view so that we can re-use it whenever it pleases us.

However, suddenly we realize that there is a potential problem with our query:
As we discussed before, customer names are not necessarily \sqlilIdx{UNIQUE} in our \db\ design.
Product names are \sqlil{UNIQUE}, but eventually, we may have two customers with the same name.
This would be a flaw of this query, as it does not give us any means to distinguish two different customers with the same name.
In real systems, you have customer numbers, unique usernames, or phone numbers to avoid this issue.
Actually, we \emph{do} have phone numbers here, too, and they \emph{are} \sqlilIdx{UNIQUE}.
Then again, just displaying the phone number is probably confusing to whoever reads the result of the view.

So what are we going to do?
Shall we display the customer names, which are not unique?
Then, if we add up sales per customer \emph{name}, people with the same name will be grouped together and the result may be wrong.
Or should we display phone numbers?
This would render the output close to unreadable for human operators.%
%
\begin{sloppypar}%
Let's split the difference:
We simply concatenate the customer name and their phone number.
We not just select \sqlil{customer.name AS customer_name}.
Instead, we write \sqlil{customer.name || ', ' || customer.phone AS customer_name}.
The \sqlil{a || b}\sqlIdx{\textbar\textbar} operator concatenates two strings \sqlil{a} and \sqlil{b}~\cite{PGDG:PD:SFAO}.
Thus \sqlil{'Hello ' || 'Word!'} will result in the string \sqlil{'Hello World!'}.
Our third customer, Mrs.~Bebba, would be displayed as \sqlil{'Bebba, 33333333333'}.
Since the phone numbers are \sqlilIdx{UNIQUE} in our \db, customer names plus phone numbers must also necessarily be unique.
As the result, the data is readable, because the human operator can see the names.
And if ever another Mrs.~Bebba would become our customer, maybe she has phone number~77777777777, then it would be impossible to mix up the two Bebbas.
One would show up as \sqlil{'Bebba, 33333333333'} and the other one as \sqlil{'Bebba, 77777777777'}.%
\end{sloppypar}%
%
We now can confidently save this improved query into the \db\ as a \emph{view}.
This way, we can re-use it whenever we want.
Storing it as a view is very very simple.
We first need to pick a name for it.
Let's call it \sqlil{sale}.
Then, we just need to write \sqlil{CREATE VIEW sale AS}\sqlIdx{CREATE!VIEW}\sqlIdx{VIEW}\sqlIdx{AS} right before our new query~\cite{PGDG:PD:CV}.
If we do this, the query is not actually executed.
It is stored under the name \sqlil{sale} as a view.
We do this in \cref{lst:factory:create_view_sale}.

Then how do we actually execute the query?
Simple:~We can treat \sqlil{sale} as if it was a table.
We write \sqlil{SELECT * FROM sale;}.
We do this as well in \cref{lst:factory:create_view_sale}.

The result is shown in \cref{exec:factory:create_view_sale}.
Eight beautiful human readable rows of data.%
%
\FloatBarrier%
\endhsection%
%
\hsection{Using our View}%
%
\gitLoadAndExecSQL{factory:select_from_view_sale_1}{}{factory}{select_from_view_sale_1.sql}{factory}{boss}{superboss123}%
\listingSQLandOutput{factory:select_from_view_sale_1}{%
Compute the per-customer and per-product sales based on the view \textil{sale}.%
}{}%
%
\gitLoadAndExecSQL{factory:select_from_view_sale_2}{}{factory}{select_from_view_sale_2.sql}{factory}{boss}{superboss123}%
\listingSQLandOutput{factory:select_from_view_sale_2}{%
Compute the per-customer and per-product sales based on the view \textil{sale}.%
}{}%
%
Above, we said that the view \sqlil{sale} can be used like a table, with the exception that you cannot add or change data.
\sqlil{SELECT * FROM sale;} looks very much like that.
But we can do a lot more.
We now want to answer the questions \inQuotes{Which customer bought products for the most money for us?} and \inQuotes{Which product sold for the most money?}

The data we need to answer these questions is already available in the view \sqlil{sale}.
For each demand, we have the \sqlil{price} of one unit of the ordered product as well as the \sqlil{amount}, the number of product units purchased, in our table.
If we multiply them with each other, i.e., compute \sqlil{amount * price} we know how much the customer paid.
All we need to do is to sum up this quantity per customer and we know how much money each customer sent to us.
We can also sum \sqlil{amount * price} for each product and we get how much money came in on a per-product basis.%
%
\begin{sloppypar}%
In the former case, we would \sqlil{GROUP BY customer_name} and then \sqlil{SELECT customer_name, SUM(amount * price)}\sqlil{SUM} from \sqlil{sale}.
We could give the sum the name \sqlil{customer_sale}.
To directly see who purchased most from us, we could also add a \sqlil{ORDER BY customer_sale DESC}\sqlIdx{ORDER BY!DESC}.
The \sqlil{DESC} enforces sorting in descending order, i.e., larger values come first.%
\end{sloppypar}%
%
We can do pretty much the same for products, in which case we would \sqlil{GROUP BY product_name}.
We could compute \sqlil{SUM(amount * price) AS product_sale}\sqlIdx{SUM} and we would know for how much money each product sold~\cite{PGDG:PD:AF}.
On a per-product basis, it makes also sense to look at the total units sold, so we could additionally compute \sqlil{SUM(amount) AS total_amount}\sqlIdx{SUM}.
Of course, we could sort the data by the product sale in descending order, i.e., write \sqlil{ORDER BY product_sale DESC}\sqlIdx{ORDER BY!DESC}.

The complete queries can be seen in \cref{lst:factory:select_from_view_sale_1}.
Their result is given in \cref{exec:factory:select_from_view_sale_1}.
We find that Bebbo is our most valuable customer.
He purchased products for over 2996Ԫ from us.
Bebba purchased for over 2570Ԫ and Bibbo sent about 2433Ԫ to us.

The top-selling product, by far, is our famous \inQuotes{Shoe, Size 42.}
We sold 12 pairs of it for about 1956Ԫ.
Figuratively speaking, \inQuotes{Shoe, Size 42} is the runner up~(hehe) with 1590Ԫ for ten pairs sold.
Many of our customers also liked to by \inQuotes{Large Purse.}
No wonder that customers love it, as it offers 10g of product per~Ԫ~(see \cref{exec:factory:select_from_table_product}).

It is interesting to briefly think about how the above works.
On the top level, we have an \sql\ query that selects from data from the view \sqlil{sale}.
Fine, you already have a feeling about how such queries work.
But remember:~\sqlil{sale} is \emph{not} a table.
It itself is just a query!
You can imagine that what happens here is that the query stored as view \sqlil{sale} is executed.
Its results are taken and then on top of these results, the new query that sums up the customer purchases~(or the product purchases) is executed.
The results of \sqlil{sale} are not stored.
\sqlil{sale} is not a table.
Instead, whenever we access it, the query it represents is executed.
And we can work with its results as if it was a table.
Isn't that cool?

Assume that your \db\ is now used for several years.
Queries like the above one would take into consideration the whole history of data.
Maybe we are only interested in the present year, which, at the time of this writing, is~2025.
So what we would like to do is to limit our query from \cref{lst:factory:select_from_view_sale_1} to only consider data from~2025.%
%
\begin{sloppypar}%
For this, we add the clause \sqlil{WHERE (ordered >= '2025-01-01') AND (ordered < '2026-01-01')}\sqlIdx{WHERE}\sqlIdx{AND} to our queries.
\sqlil{ordered} is the date when the customer made the order.
We only want to consider records for which the date is greater then or equal to January~1st, 2025.
Also, the date must be less than January~1st, 2026.
So we combine \sqlil{(ordered >= '2025-01-01')} and \sqlil{(ordered < '2026-01-01')} with \sqlilIdx{AND} and put them into a \sqlilIdx{WHERE} clause.
The updated queries can be found in \cref{lst:factory:select_from_view_sale_2}.%
\end{sloppypar}%
%
Their results in \cref{exec:factory:select_from_view_sale_2} show us that Mr,~Bebbo is still our most valuable customer in 2025, and that Mrs.~Bibba still bought products for more money than Mr.~Bibbo.
However, in 2025, \inQuotes{Large Purse} is our best-selling product with a margin of about 600Ԫ over the runner up.
\inQuotes{Shoe, Size 42} did not even make the list in~2025.

At this stage, we have again become more powerful.
We now have effectively solved the problem that our data is hard to understand.
We now have tools, namely the \sqlilIdx{JOIN} queries, that allow us to pull together the data from different tables.
We originally separated the data to enforce cleanliness without redundancy.
We stored the information about Mr.~Bebbo only once, even though he made four purchases.
The downside was that, in the purchase records for Mr.~Bebbo, the name \inQuotes{Bebbo} did not appear.
Instead, it was referenced by pointing to another table via its id~2.
This made the data hard to read.
But now we know how we can put the records back together.

Admittedly, the sales query that pulls all the data together~(given in \cref{lst:factory:create_view_sale}) is not very easy to read if you are new to \sql.
On one hand, once you are more familiar with \sql, such queries become a joy to write.
On the other hand, we also can simply store the query in the \db\ as a so-called \emph{view}.
We can re-use this view whenever we want.
We can also plug other queries on top of it.
So we basically have to do the hard work of designing a good query to pull the data together only once.%
%
\FloatBarrier%
\endhsection%
\endhsection%
%
