%
\hsection{Creating a User and the Database}%
%
\gitSQLAndOutput{\databasesCodeRepo}{factory}{create_user.sql}{}{}{}{postgres.sh}{factory:create_user}{%
Using \sql\ to create a the user \textil{boss} with password \textil{superboss123}.%
}%
%
The first step to fulfill this request to is to create a new and empty \db.
We already installed \postgresql\ (see \cref{sec:installingPostgres}).
It is running on a dedicated \server\ computer in our small \pgls{IT} department / office.
This \db\ \server\ will host all the \dbs\ of the company, probably ranging from payroll data to fancy business analytics.

Our boss wants to have full access to the new \db.
They are not \pglspl{dba}, though, so we would only reluctantly give them administrative access to the complete \server.
In a first step, we would like to create a new role or user account on our \dbms.
This account should only be able to access the new factory \db.
If they make a mistake, this mistake will only affect this single \db.
If some outside attacker can obtain their password, then the impact will only be limited to this \db\ and not affect, e.g., payroll data or other confidential data in other \dbs.

For the sake of simplicity, assume that we are logged in the \db~\server\ computer and that the password to the administrator user \textil{postgres} is set to~\textil{XXX}.
We open a \pgls{terminal}, which can be done via \ubuntuTerminal\ under \ubuntu\ \linux; under \windows, \windowsTerminal.
We wanot to start the \client\ program \psql\ with the proper connection \pgls{URI}~\cite{PGDG:PD} to access our \postgresql\ \server.
We can connect to the \postgresql\ \server\ on our local machine by providing the connection \pgls{URI} \textil{postgres://postgres:XXX@localhost}.
The \textil{postgress://} tells \psql\ that this is, indeed, a connection \pgls{URI}.
The second \textil{postgres} is the user name.
The colon~\inQuotes{\textil{:}} separates the user name from the password~\textil{XXX}.
\localhost\ stands for the current machine itself.

Let me stress again:
Never use something like \textil{XXX} as a password.
I am also not doing that.
For the examples in the book, I just replaced the actual password with \textil{XXX}.

Anyway, after the \inQuotes{\textil{@}} comes the network address or host name of the where the \postgresql\ \server\ is running.
Since it is on the same machine as we are currently working on, we write \localhost.
With this information, \psql\ knows that we want to connect to the \postgresql\ \server\ running on our current computer as user \textil{postgres} with password~\textil{XXX}.%
%
\begin{sloppypar}%
Once the \psql\ \pgls{terminal} is open, we can begin typing commands in \sql.
We want to create a new user for the \psql\ \server.
As username, we pick \inQuotes{boss}.
The password shall be \inQuotes{superboss123}.
Obviously, this is a very unsafe password.
The boss will have to change it as soon as possible.
Anyway, in order to create the new user with that password, we write \sqlil{CREATE USER boss WITH ENCRYPTED PASSWORD 'superboss123';}.
The first part of the command, \sqlIdx{CREATE!USER}\sqlil{CREATE USER xyz}, tells the \server\ to create a new user account under the name \textil{xyz}.
The second part, \sqlil{WITH ENCRYPTED PASSWORD 'abc'}\sqlIdx{WITH!ENCRYPTED PASSWORD}, tells the \server\ that the password \textil{abc} should be used for this user.
Passwords are stored in an encrypted way anyway in \postgresql, but it never hurts to specify this clearly.
Maybe we want to run the same commands later on another \dbms\ where it is necessary to explicitly say that passwords shall be encrypted.
You can read more about \sql\ commands in the \postgresql\ reference~\cite{PGDG:PD:SC}.%
\end{sloppypar}%
%
\bestPractice{keywords}{%
Keywords in \sql\ may always be written completely in uppercase~\cite{DB2019HTTPSSC}. %
Well, technically, \sql\ keywords are not case sensitive, so \sqlilIdx{WHERE} and \sqlil{where} work the same. %
It is most important to be consistent in your casing, regardless whether you prefer upper- or lowercase~\cite{B2025SBPASG}. %
Nevertheless, I prefer uppercase and the \postgresql\ documentation does so too~\cite{PGDG:PD}.%
}%
%
Anyway.
How do we know that the user \inQuotes{boss} has successfully been created?
On one hand, \psql\ did not give us an error message.
On the other hand, we can simply list all users.
See, in \postgresql, all users are stored in the table~\sqlilIdx{pg\_catalog.pg\_user}.\footnote{%
On other \pglspl{dbms}, the users may be stored differently.}
So, we can just query this table.
The \sql\sqlIdx{SELECT{\idxdots}FROM} command \sqlil{SELECT usename FROM pg_catalog.pg_user;} will list the value of the column \textil{usename} for all rows in the table \sqlilIdx{pg\_catalog.pg\_user}.
If we run this command \emph{before} creating the user, on a fresh \postgresql\ installation, it will only list the single user \textil{postgres}, i.e., the administrator of the whole \dbms.
If we run it again \emph{after} creating the new user, it will also list~\textil{boss}.

The complete steps of querying the existing users, adding the new user \inQuotes{boss}, and then querying the list of users again is shown in \cref{lst:factory:create_user}.
The output that we will get from running this small script is given in \cref{exec:factory:create_user}.
Notice that you can download the whole example and run exactly the commands (the {\color{listing-tool-command}{red text}} after the {\color{listing-tool-command}\texttt{\$}}) from our repository \url{\databasesCodeRepo}.

\gitSQLAndOutput{\databasesCodeRepo}{factory}{create_database.sql}{}{}{}{postgres.sh}{factory:create_database}{%
Using \sql\ to create a database for user \textil{boss}.%
}%
%
\begin{sloppypar}%
Having created the new user~\inQuotes{boss}, we can now create the \db\ \inQuotes{factory} to be owned and worked on by that user.
For this, all we have to do is to type the \sql\ command \sqlil{CREATE DATABASE factory OWNER boss;}\sqlIdx{CREATE!DATABASE}\sqlIdx{OWNER}.
This command pretty much explains itself.
It will create a new \db\ with the name \textil{factory}.
The user \textil{boss} will be owner of this \db, i.e., they will have full access to add and manipulate its data.
Notice that we still need to execute this command under the \pgls{dba} role \textil{postgres}.
This user has the right to create \dbs\ for other users which can then work with them.%
\end{sloppypar}%
%
How do we know whether the command finished successfully?
The first hint is, again, the no error message appears.
Second, we can also list all the \dbs\ in our system.
For this purpose, we write \sqlil{SELECT datname FROM pg_database;}\sqlIdx{SELECT{\idxdots}FROM}\sqlIdx{pg\_database}.
See, all the names of all \dbs\ inside the \postgresql\ \dbms\ are stored in the column~\sqlilIdx{datname} of the table~\sqlilIdx{pg\_database}.\footnote{%
On other \pglspl{dbms}, the \pglspl{db} may be stored differently.}
If we run this command \emph{before} creating the new \db\ on a fresh \postgresql\ installation, we find that it will list some standard \dbs, which we will ignore here.
Running the command again \emph{after} creating the new \db\ now lists one more \db:~\textil{factory}.
The steps of this example are given in \cref{lst:factory:create_database} and the output of the \psql\ \client\ is given in~\cref{exec:factory:create_database}.%
%
\FloatBarrier%
\endhsection%
%
