\hsection{Further Reading}%
\begin{noglslink}%
%
The author of this book is not really a \db\ buff.
I did several \db\ projects, some for hobby an some even in commercial use, but neither of them had a really large scale.
And most of the things I did were self-taught, mainly through technical manuals and documentation, complemented by classes I took at the university.
This book is an attempt to create a resource that can help students to learn about the field of \pglspl{db} on a practical level, with links to theory.

There are many resources, internet websites as well as physical books, that are of very high quality.
Actually, this book cites and builds upon several of these resources.
They complement, extend, and expand upon the topics that we discuss here.
For every fact that we explain in this book, we try to cite proper sources at the end the book.
Please always feel free to supplement the reading of this book with the following resources.%
%
\hsection{Lectures at Universities}%
Here we provide a list of lectures at different universities for which the slides are online available.
When we discuss some topics, where appropriate, we will also reference the corresponding slides of these classes as easily-accessible and concise further reading.%
\begin{itemize}%
\item \furtherReading{S2024D},%
\item \furtherReading{KC2024DS},%
\item \furtherReading{P2024C6DS},%
\item \furtherReading{T2025CDBMS},%
\item \furtherReading{C2016CDMS},%
\item \furtherReading{R2024CDS},%
\item \furtherReading{G2011EW2ITDS} as well as \furtherReading{G2011ES2DSI},%
\item \furtherReading{P2006CITRD},%
\item \furtherReading{SS2005EIDDDFDB}, and%
\item \furtherReading{V1999C5DMS}.%
\end{itemize}%
\endhsection%
%
\hsection{Books}%
The following books may be useful for further reading:%
\begin{itemize}%
\item \furtherReading{PW2022APITD},%
\item \furtherReading{SKS2019DSC},%
\item \furtherReading{HRT2021MDM},%
\item \furtherReading{EN2015FODS},%
\item \furtherReading{KE2015DEE},%
\item \furtherReading{TATPC2009TMGTD},%
\item \furtherReading{H2020DDFMM2AE},%
\item \furtherReading{K2017SAATPODP},%
\item \furtherReading{CM2018DSDIM},%
\item \furtherReading{D2003AITDS},%
\item \furtherReading{FP2023LP},%
\item \furtherReading{RG2002DMS},%
\item \furtherReading{H2016RDDAI},%
\item \furtherReading{GMUW2008DSTCB},%
\item \furtherReading{HM2024IMARD},%
\item \furtherReading{BHS2015RIDS},%
\item \furtherReading{MS2001S1URLC},%
\item \furtherReading{CM2000MDMAUDA},%
\item \furtherReading{T2018ISARD},%
\item \furtherReading{SP2002STYSI2D} and its German translation~\furtherReading{SPMP1998SI2TDDASVEI12T},%
\item \furtherReading{MM2014RDAMA}, and%
\item \furtherReading{TA2024DDAMWPAM}.%
\end{itemize}%
\endhsection%
%
\hsection{Websites}%
Some websites that you might find useful are:%
\begin{itemize}%
\item \furtherReading{PGDG:PD},%
\item \furtherReading{SE:DA},%
\item \furtherReading{B2025DS}, and%
\item \furtherReading{D2022DN}.%
\end{itemize}%
\endhsection%
%
\end{noglslink}%
\endhsection%
%
