\hsection{Requirements Analysis}%
%
The requirements analysis is one of the most important steps of the \db\ development lifecycle.
It is this point where we gain the understanding of the project.
Requirements engineering mediates between the users and customers of a project on one side and the developers and suppliers on the other~\cite{ISOIECIEEE2018SASELCPRE}.
The results of the requirements analysis are \pgls{SRS} documents that~\cite{ISOIECIEEE2018SASELCPRE}%
\begin{itemize}[noitemsep]%
\item enable an agreed understanding of all stakeholders~(acquirers, users, customers, operators, developers, and suppliers),%
\item have been validated against the real-world needs,%
\item can be implemented, and%
\item provide a reference for verifying designs and solutions.%
\end{itemize}%
%
On one hand, during the requirements analysis, we build a clear understanding of the purpose, goals, and limits of the project.
On the other hand, we also need to learn about the organizational structure and processes that should be embodied by our \db\ and the applications on top of it.
Indeed, studies show that for many companies, more than half of the problems of systems~\cite{S2003ISA6P:RDARS} and costs of software development~\cite{IC2009BAB2TPTS} are based on poor requirements definition.
Inefficient requirements management is considered a top-cause for project failure~\cite{EDN2005RUIFACI}.
Good requirements engineering can increase the developer productivity and lead to improved project planning~\cite{DCVP2005READSDFFAC}.
Fixing errors in the requirement cost 10 to 200 times as much once the application is deployed compared to discovering them during the requirements analysis~\cite{BP1988UACSC,M2001FTEAOOP,RGJ2023EASARBFSAR}.
The collection and analysis of requirements therefore is very important.%
%
\hsection{Types of Requirements}
Requirements can roughly be divided into business requirements, functional requirements, non-functional requirements, and constraints~\cite{I2018SAH}.

First, the business requirements are the high-level goals, objectives, and sought outcomes of the project.
They define the motivation of the organization for why the system is being developed~\cite{ISOIECIEEE2018SASELCPRE}.
An organization usually has some overall initiative or plan to improve some of its metrics.
The project is initiated to support this initiative.
These requirements therefore defined such that the project will contribute to improving the organization.
They are more general and abstract.

Second, the functional requirements are more concrete and define what the system should do.
They define the functions, features, and the behavior that the system must offer to fulfill the business requirements.
They may be defined as%
\begin{itemize}[noitemsep]%
\item the input given to the system,%
\item the expected operation to be performed by the system on that input, and%
\item the expected output to be produced.%
\end{itemize}%
%
Third, the non-functional requirements define how the system should perform.
They define the service quality that the system should offer, as well as the required the performance, usability, scalability, reliability, etc.
This includes the computational environment in which it must be possible to execute the system.

Fourth and finally, the constraints are the factors that limit which solutions are viable.
They define the boundaries in which we operate.
While requirements define properties that our system should have, the constraints rule out methods to get there.
They can appear in form of budget or time limits.
They may also appear in the form of demands like \emph{\inQuotes{Your system must work with version~XYZ of software~ABC.}}%
\endhsection%
%
\hsection{Requirements Gathering}%
%
Several different methods exist that can be used to collect requirements~\cite{S2003ISA6P:RDARS}.

The first method, interviewing stakeholders to gather information about the system that we will design, is considered one of the most efficient requirements gathering techniques~\cite{DTHJM2006EORETERDFASR,S2003ISA6P:RDARS}.
For each interview, careful preparation is needed, which includes making an appropriate appointment.
While stakeholders may spontaneously share information on some topics, they may not discuss others unless explicitly prompted with questions~\cite{BJF2014WSWOWNSATAESOTIIREEI}.
Each interview meeting should therefore have a proper agenda, pre-prepared questions, and follow a checklist~\cite{WSEG2016ETIOAKOSTCORQQ}.
If possible and if the interview partner(s) consent, then the meeting should be recorded.
The recordings and notes should be evaluated within two days after the interview.
The interviews should neutral, not push the interviewee into any direction.
The goal is to collect diverse views on the project.

A second, more standardized method to collect information are questionnaires~\cite{TAE2008ISRTDQTSTDQSSAE,S2003ISA6P:RDARS}.
This way, we can collect much data from many stakeholders within a brief time.
Questionnaires are easy to evaluate and process, but designing them properly is important.
Different types of stakeholders will use different parts of our system.
Therefore, it may be necessary to design several different questionnaires, one for each group of future users.
The questions should be clear an unambiguous.
For some aspects, multiple-choice questions or range-based ratings are good, while others may require open-ended questions where the users can fill in their opinions.

The third method is to directly observe users doing their work~\cite{SGGM2007TMSPATFISRDWS,S2003ISA6P:RDARS}.
Stakeholders may not always be able to accurately describe their function and how they fulfill it.
Observing them performing their day-to-day processes can thus provide helpful additional information.
It is also possible to record such real-world examples instead of personally observing them~\cite{HPW1998REAVWRWS}.
Of course, under observation, people may behave somewhat differently from normal, so such information is to be taken with a grain of salt.

Fourth, we can also analyze both organization-internal and external regulations, as well as published procedures, processes, and other documents~\cite{RGJ2023EASARBFSAR,S2003ISA6P:RDARS}.
Usually, an organization will have their own written regulations, announcements, and otherwise standardized procedures.
Additionally, there are regulations and laws imposed upon an organization and the processes within.
Gathering such information can be crucial and complements our understanding of what the system is supposed to do, and why current systems work the way they do.
Sometimes, the official documentation of organization-internal methods and the practical realization of processes may differ, though.

A fifth method are group meetings and workshops~\cite{S2003ISA6P:RDARS}.
Here, under the guidance of a session leader, stakeholders at all levels meet, from management to end user, from system analyst to data entry personnel.
The group then jointly discusses the current situation and the planned system, which can be a highly efficient way to gather requirements.
Different variants of this method exist since the 1970s, under names such as \glsreset{JAD}\pgls{JAD}~\cite{CWG1993PAJADATC,M1996RTWSS} and \glsreset{participatoryDesign}\pgls{participatoryDesign}~\cite{CWG1993PAJADATC,FMRSW1989OOSAATSDASD}.

Several of these techniques can be combined, often with other approaches such as brainstorming sessions, surveys, reverse engineering of existing systems, or prototyping~\cite{I2018SAH,S2003ISA6P:RDARS}.
In the context of \pglspl{db}, it is particularly important to properly define the data structures and entities when analysing the requirements and, later, when developing the conceptual model~\cite{M1987DFADSMFDRACS}.%
%
\endhsection%
%
\hsection{Requirements Specification Document}%
After the requirements have been collected, they are stored in a formal specification document, the so-called \glsreset{SRS}\pgls{SRS} document~\cite{S2010DSRSAR,W2004ASOTMFMTQOTRSD}.
The \pgls{SRS} is the most important document in the software development process.
The document structure should follow the IEEE~830\nobreakdashes-1998~\cite{IEEE1998IRPFSRS} standard or the newer ISO/IEC/IEEE~29148\nobreakdashes-2018~\cite{ISOIECIEEE2018SASELCPRE} standard.
While the lifecycle of software or systems in general can be managed by ISO/IEC/IEEE~15288~\cite{ISOIECIEEE2023SASESLCP} and ISO/IEC/IEEE~12207~\cite{ISOIECIEEE2017SASESLCP}, respectively, ISO/IEC/IEEE~29148~\cite{ISOIECIEEE2018SASELCPRE} provides the guidelines for their requirements-related processes.
Generally, it would be a good idea to simply follow these standards when gathering and analyzing requirements for software or \db\ projects.%
\endhsection%
%
\hsection{Example: Teaching Management Platform}%
%
We will now explore the gathering and analyzing of requirements by using an example.
Our example project is the development of a teaching management platform for a university.
Sadly, doing this at a realistic level would go far beyond what we can do as a reasonable example in the context of a course book.
We cannot really implement complete outline of ISO/IEC/IEEE~29148\nobreakdashes-2018~\cite{ISOIECIEEE2018SASELCPRE}.
We cannot even specify the complete and exact requirements of any realistic system without exceeding reasonable time and length limits.
Therefore, we will try to discuss the requirements partially.
We will pick some more or less interesting issues while leaving others to the imagination of the reader.
Also, we write down everything in a very informal way.
In actual documents, a much more formal language would be used.%
%
\hsection{Business Requirements}%
%
Our imaginary university has several goals that it wants to achieve by introducing a new teaching management platform.
At the start of our project, we had several meetings with the university leadership.
During these meetingsm we gathered and understood the business purpose of the project.
We also get a good idea about the roles of the stakeholders involved.%
%
\hsection{Business Purpose}%
First, by migrating processes into a \db\ and designing applications to access them via the web, all stakeholders will benefit:
The students can more easily access their courses, obtain transcripts, register for classes, etc.
The professors can manage their classes more efficiently.
The workload of the management of the university can significantly be reduced and simplified.

Second, the administrative personnel so far does all the management of its students and courses using \microsoftExcel\ sheets and pen and paper.
This has several drawbacks, such as the lack of centrally controlled backups, the possibility of errors, the lack of traceability of processes, problems when this duty is eventually handed over from one teacher to another, and so on.
By developing a centralized system, the imaginary university wants increase the control over as well the accountability, traceability, and documentation of its processes.
This could be seen as a tool for supporting quality management, maybe along the lines of ISO~9001~\cite{ISO180912019,ISO90012015}.

Third, the long term goal is the digital transformation of our whole imaginary university.
All processes inside the imaginary university would be managed by online platforms.
This would significantly reduce administrative efforts and costs.
All processes would automatically be documented and backed up.
The quality of the services rendered to students would improve.
Auditing becomes easier.
The teaching management platform is the first building block of this digital transformation.
It will allow the university to gather experience with systems that are not handled just by a few administrative personnel~(such as the \pgls{HR} or financial accounting system), but accessed by thousands of users in different roles.%
%
\endhsection%
%
\hsection{Major Stakeholders}%
%
There are five main groups of stakeholders on the university side of our imaginary university management system.

First, there are the students.
The students need to register online for modules~(courses and exercises).
They want to print their schedule, which includes which courses they will attend at which days and times and in which rooms.
They want to view their scores and progress.

Second, there are the faculty members, i.e., the professors and teachers.
A professor can chair a module, meaning that they will teach a certain class.
Other faculty members, say lecturers or assistant professors, may teach lab classes or practice classes.
They have access to the list of students in their respective modules.
Like students, teachers can see their schedule.

Third, the university administration can create and manage schools in the system.
The university administration also manages the professors, teachers, and students.
They are the only ones who can enter new persons into the system.
They are the only ones who can create new curricula for the different schools in the system.
They assign buildings and rooms to schools.

Fourth, there is the administration of the different schools of the university.
The school administration assigns professors to modules, lecturers to practice classes.
They are allowed to create and schedule exams and major deliverables.
They manage their rooms.
They manage the students belonging to their school.

Fifth and finally, there are system administrators, including the \pglspl{dba}.
Their most important task is to keep the system running.
This means that they will run regular backups.
They also need to practice regularly how to restore the system from the backups.
They also need to update all the involved components, such as the \pgls{OS}, the \dbms, the web \pglspl{server}, and so on.%
\endhsection%
%
\endhsection%
%
\hsection{Functional Requirements}%
After the stakeholders and motivation of the project have been reasonably clarified, we begin by interviewing several involved personnel.
We visit the education department of the university.
We interview the deans, vice deans for teaching, and secretaries of three different schools.
We discuss with five professors.
We also meet with the students union and several students at different academic performance levels.
Our goal is to understand the academic processes from the points of view of different sides.

We want to clarify several questions.
What is the basic functionality that we need to provide?
What features could we offer that, currently, are unavailable?
Based on our findings, we create different questionnaires and distribute them more widely among the above peer groups.
Finally, we collect all the information and present them in workshops where, again, members of all the above groups take part.
It is our goal to build a view on the requirements that can be agreed to by all stakeholders.
We then continue writing our \pgls{SRS} document as follows.

The system has to be available through one or multiple websites.
Several processes must be supported, for example:%
%
\hsection{Person Management}%
Only the university administration can create student or faculty member records in the system.
Such a record must store information such as name, ID, mobile phone number, gender, highest academic degree, academic rank, role~(student, staff, \dots)etc.
The university administration must be able to change these information.
They can assign the people to schools.
The administration of a school can access only the data of the people assigned to them.%
\endhsection%
%
\hsection{Date Management}%
The university administration sets dates such as semester begin and end, begin and end of the exam period, holidays, etc.
The university administration provides ranges for special dates such as the start of a graduation projects~(开题), the graduation project midterm evaluations~(中期), or the graduation project defenses~(毕业).
The schools then can adapt these date ranges to their situation.
Professors who chair modules can then, for example, schedule an exam in the time range defined by their school, which, in turn, is a sub-range of the exam schedule provided by the university.%
\endhsection%
%
\hsection{Curriculum Management}%
The university administration can create and change curricula.
They do this of course in cooperation with the schools, but only they have the right to make decisions.
Each curriculum contains different modules at different semesters.
Modules can be compulsory or optional and belong to certain types, such as Subject Basic Courses~(学科基础课),
General Basic Courses~(公共基础课),
Professional Basic Courses~(专业基础课),
Professional Elective Courses~(专业选修课), or
General Elective Courses~(公共选修课) and such and such.
Modules can consist of only lectures, of lectures and practical training, only practical training, or deliverables~(such as BSc and MSc theses).
The university administration then assigns curricula to schools.
The school administration can enroll their students to curricula.%
\endhsection%
%
\hsection{Module Management}%
The school administration can, in each semester, create implementations of the modules.
An implementation assigns a teacher to course and has an upper limit for student enrollment.
Teachers are notified about their assigned courses.
Teachers can print their teaching schedule.%
\endhsection%
%
\hsection{Room Management}%
The university administration manages, creates, changes, or deletes records for lecture rooms.
Rooms have locations in buildings, a capacity for students, and features such as equipment~(overhead projectors, blackboards, lab equipment for computer science, chemistry, biology, \dots).
The university administration can assign rooms to schools.
The school administration can assign courses to rooms and timeslots.%
\endhsection%
%
\hsection{Module Enrollment}%
Students can enroll to the modules in their respective curricula.
For compulsory modules that are offered only by a single professor, they are automatically assigned.
Otherwise, the school administration can manually assign them.
For any module in their curriculum for which they are not assigned by the school, they can choose by themselves.
Students are notified automatically about enrollment options.
Students can print their schedule.%
\endhsection%
%
\hsection{Exams and Deliverables}%
The school administration can create exams for each module.
An exam has an assigned room and time in the exam period.
Professors can also request the school to schedule midterm exams, if they want.
A midterm exam has an assigned room and time outside the exam period.
For each module they teach, professors can also create deliverable records, for example for homework.
They can then assign scores for the students for the exams and deliverables.
Students are automatically notified about exam dates and locations for their modules.
Students can print their current transcripts as well as the scores of all deliverables for each module they take at any time.
\endhsection%
%
\hsection{Communication}%
The system offers a facility for communication between students, teachers, and their school.
The communication records are immutable and will be preserved.
While this channel is likely not used often, it may be useful for things such as reminders, notifications, but also warnings or objections.%
\endhsection%
%
\hsection{Administration and Backup}%
The \pglspl{dba} of the university can create backups of the platform.
They can update the platform.
There always is a second instance of the platform to test updates and to verify that the backup procedure is working.
The \pglspl{dba} can also install the platform on a new computer and load the backups.%
\endhsection%
\endhsection%
%
\hsection{Non-Functional Requirements}%
The websites must render correctly and be usable both on desktop computers as well as mobile phones.
They must work under the out-of-the-box default web browsers provided by \microsoftWindows, \linux, \macOS, \appleIOS, \iPadOS, and \android.

The system must handle 50\decSep000~new student records, 2000~new staff records, 200~curricula, 8000~modules, and 1\decSep000\decSep000~new exam/deliverable results per year.
It must be able to handle this load over 20~years, i.e., twenty times the above.
After twenty years, we assume that either the hardware is upgraded or that old records are removed from the system and backed up elsewhere.
The response time for any query of any application must never exceed 2~seconds.

Communication should be secured over \pgls{HTTPS}.
Proper data protection must be offered, i.e., people can only access data relevant to them and this access is secured.%
\endhsection%
%
\hsection{Constraints}%
The system must be set up as a set of \docker\ containers in the computational center of the university.
The system should be composed entirely of open source software, in order to increase reliability, availability, and to reduce costs.
The computational center of the university will provide three computational nodes, each with a 32~core processor with 4~GHz clock speed and 64~GiB of RAM.
The computational center of the university will provide 100~TiB of centralized storage space.
A first prototype must be developed within six months.
The system must enter thorough test and validation within one year.
The project budget is limited to 5\decSep000\decSep000~RMB.%
\endhsection%
%
\hsection{Summary}%
From this requirements analysis, we learned a lot.
We understand that this is actually quite a complicated system.
We also see that it does involve a lot of different aspects.
Many \db\ projects do.

Like in many \db\ projects, there is not just the \db.
There also are web-based \glsreset{GUI}\pglspl{GUI}.
Our course and book, however, are not about user interface design.
We will just assume that we are part of a team here, and somebody else will take care of the \pgls{GUI}.
We also do not really care about the \docker\ software environment structure either.
We also just assume that the budget, the other constraints, and the non-functional requirements are OK.
In the following, we will focus entirely on the \db\ part of this project.

This project would be immensely more complicated than our simple example from back in \dref{sec:simpleExampleFactory}.
However, if you think back on what we have already touched, \sql, forms, and reports {\dots} then you may have some rough idea on how this project here could be tackled.
OK, the tables that we are going to need will be more complicated.
But they will still be tables.
The \sqlilIdx{JOIN} and \sqlilIdx{REFERENCES} keywords will still work and be our basic tools to connect and merge data.

Of course, the system should have a web-based \pgls{GUI} facing the students -- but, as said, we will ignore this and assume that this can somehow be done via \python.
At least we can imagine to use \libreofficeBase\ to construct at least some raw prototype for the administrative staff, where data could be entered and where schedules could be printed as reports.

Clearly, this undertaking will require much more brain power from our side.
But if we tackle this in a well-structured way step-by-step {\dots} then maybe we do have a decent chance.
Actually, doing things in an well-organized {\dots} that's all what we have been talking about recently.
First we begin with a conceptual model, then we derive a logical model {\dots} that's exactly this kind of stuff.
Let's do it like this!%
\endhsection%
%
\endhsection%
\endhsection%
%
