%
\hsection{Getting Started}%
%
The goal of this course is to introduce databases.
However, this is a practical course.
So we will work on practical examples and create (more or less) \inQuotes{realistic} \pglspl{db} on real \pglspl{dbms}.
This means that we will look into several interesting topics.
And for most topics, we will try to get practical hands-on experience using a suitable software tool.%
%
\begin{itemize}%
%
\item We will work with an actual \pgls{dbms}.
So you need to install an actual \pgls{dbms}.
We choose \postgresql~\cite{TA2024DDAMWPAM,FP2023LP,OH2017PUAR,B2024PELUYDW}.%
%
\item \pglspl{dbms} are usually just \pglspl{server} to which you can send \sql\ commands via a \pgls{client} \pgls{terminal}.
However, there exist also nicer tools offering rich \pglspl{GUI} that allow you to design forms and reports.
We will try using such a tool as well.
We choose \libreofficeBase~\cite{FNFHWSKLSSGLFRSRPLJG2022BG7R1BOL7C,S2022L7PFEUU}.%
%
\item The \pgls{dbms} is usually just the backend of a landscape of tools in an enterprise or organization.
Rarely will users work only with or directly on a \pgls{dbms}.
Instead, there may be several applications that connect to \pglspl{db} through unified \pglspl{API}.
Since we will also take a look at how that works, we need to install a programming language and corresponding \inQuotes{\pgls{dbms}-access library.}
We chose the \python\ programming language~\cite{K2018EIPFEUU,A2002PC,H2023ABGTP3P,LH2015DSAAWP}, because we also provide a free book on it in~\cite{programmingWithPython}.
As library to connect to the \postgresql\ \pgls{dbms}, we pick \psycopg~\cite{VDGE2010P}.%
%
\item An important step of the \pglspl{db} design process is to create an abstract conceptual model of the problem domain.
This involves drawing graphical diagrams, so-called \glsreset{ERD}\pglspl{ERD}, describing the real-world entities and their relationships that should be modeled in the \db.
The conceptual schema should be independent from any specific \dbms\ technology.
As tool for drawing such diagrams, we will use \yEd~\cite{SG2015MDAWY,Y2011YGEM}.%
%
\item Conceptual models need to eventually be mapped to logical models that are based on certain data models and \db\ technologies.
Interestingly, technology-specific conceptual models can also be drawn as \pglspl{ERD}, which, due to their technology-bound nature, can be translated to commands for a \dbms.
In my opinion, the best open source and free tool for drawing logical models for \postgresql\ \pglspl{db} is \pgmodeler~\cite{AES2006PPDM}.%
%
\end{itemize}%
%
\bestPractice{manyTools}{%
A computer science professional is able and always keen to learn new tools. %
A computer science professional should know dozens of different software tools for different tasks. %
A software engineer is a craftsperson and their knowledge of software is their tool belt.%
}%
%
All the tools that we consider here are free and, ideally, open source, software.
Learning about \pglspl{db} with this book should be free.%
%
\noviceHint{%
In this part of the book, we have collected installation instructions for all the tools that we use in those book.
{\color{red}{\textbf{You do not need to install all of them right away.}}}
While it makes sense to install the \postgresql\ \dbms, \libreoffice, \python, and \psycopg\ right away, because we are going to need them in the next part of the book, \yEd\ and \pgmodeler\ can wait.
Then again, maybe you want to skip over the rest of this book part and first learn something about \pglspl{db} before putting some hours of time into installing software.
Just be aware that we provide installation instructions for the tools that we need here.
Also, when we eventually need the specific software tools at some time later on, we will refer back to this part.%
}%
%
Either way, before we get into the necessary installation and setup steps for the software that we eventually need to really learn about \pglspl{db}, we face a small problem:
Today, devices with many different \pgls{OS} are available.
For each \pgls{OS}, the installation steps and software availability may be different, so I cannot possibly cover them all.
Personally, I strongly recommend using \linux~\cite{T1999TLE,B2022ELATCL,H2022LML} for programming, work, and research.
If you are a student of computer science or any related field, then it is my personal opinion that you should get familiar with this operating system.
%
\bestPractice{knowLinux}{%
Any professional computer scientist, software developer, software architect, \pgls{dba}, or system administrator should be familiar with the \linux\ \pgls{OS}.%
}%
%
Maybe you could start with the very easy-to-use \ubuntu\ \linux~\cite{CN2020ULB,H2020ULU2E}.
It should also be clear that the instructions provided here will eventually be outdated.
I am not sure whether I will be able to keep updating them in the future.
However, even a few years down the road, they should still provide some basic guidance.
In the following, I will try to provide examples and instructions for both \ubuntu\ and the commercial \microsoftWindows~\cite{B2023W1IO} \pgls{OS}.%
%
\hinput{installingPostgres}{installingPostgres.tex}%
\hinput{installingLibreOffice}{installingLibreOffice.tex}%
\hinput{installingPythonEtc}{installingPythonEtc.tex}%
\hinput{installingYed}{installingYed.tex}%
\hinput{installingPgModeler}{installingPgModeler.tex}%
%
\endhsection%
%
