\label{sec:simpleExampleIntro}%
Before we step-by-step learn about the features and intricacies of \pglspl{db}, let us look into a simple self-contained example as a teaser.
So far, you have learned about the history of \pglspl{db}.
Many courses that I know then, from here on, focus on several quite useful things:
First, an outline of the \db\ design process is given.
Second, you will learn about modeling data, how to draw \pglspl{ERD}.
Third, you will learn about normalization of data.
Fourth, you will learn the so-called $\sigma$\nobreakdashes-algebra, which is an mathematical notation abstracting from the technological aspects of \pglspl{rdb}.
Fifth, they will tell you how to select, insert, and remove data using~\sql.
Some will give really practical examples and let you explore how to work with a \db\ via homeworks.
Sixth and finally, you are taught how \pglspl{db} work internally, what datastructures they use, and how they achieve efficient storage and high query performance.

I agree that it is a good didactic method to approach \pglspl{db} from several different angles as an abstract subject.
However, when I learned something as a student, I never really learned it this way.
I learned it by doing it.
I am a believer that practical ability is nine tenth of mastering a field.
And practical ability comes easiest by playing around with the basic tools.

Also, as far as I can see, quite a few courses seem to treat \pglspl{db} as something \inQuotes{single,} something that exists \inQuotes{separately} from other things.
But this is not necessarily true.
Many of the subjects you learn elsewhere may be connected to \pglspl{db}.
For example, you often have programming classes, where you learn a programming language like \python~\cite{programmingWithPython}.
What is the connection here?
You will learn that, yeah, \pglspl{db} are often accessed from program code, maybe by so-called \emph{application servers}, that implement the business logic of an enterprise.
How does that work?
Often, you will not learn that.

Then again, \pglspl{db} are also often accessed from \pglspl{GUI}.
They maybe allow us to enter information via convenient forms.
Because users like office workers would feel puzzled if asked to enter and retrieve information using \sql\ queries.
Maybe we can also print reports with information extracted from the data in the \db.

So far, for you, a \db\ is just a nondescript thing to store data.
But you may have no idea about all the cool things that you could do with a \db.
That you could use a \db\ for your own personal purposes, ranging from keeping track of your finances over managing a bibliography of papers to storing information about your music collection.

Long story short:
We will now explore a small and self-contained example not to teach you how exactly to do things, but to show you what is possible.
And that many things are possible without in-depth knowledge and lots of work.
To make you curious.
To invite you to explore things out of your own interest and to circle back to this book to combine practical experience with background information when you like.
It is important that this is a \emph{learning by doing} example.
When reading the text, please reproduce the example step-by-step on your own computer.

Our example is a \pglspl{db} for a small company that produces shoes and handbags.
Imagine that you were hired to build an \pgls{IT}~department for the company.
On you first day, your boss enters and tells you \emph{Make an application for storing all our product variants and customer orders.}

We use a concrete technological environment for our work.
In particular, we rely on the \postgresql\ ecosystem.
We chose it because it is a very mature open source \pgls{dbms}.
It is widely adopted and was the most popular \pgls{dbms} in the \citetitle{SE:SO:2024DS}~\cite{SE:SO:2024DS} and in the open source code survey~\cite{PMPVEPWGSMB2025ATAODMSTTHOOSP}, it ranked second.%
%
\usefulTool{postgresql}{%
\postgresql~\cite{TA2024DDAMWPAM,FP2023LP,OH2017PUAR,B2024PELUYDW} is an advanced relational \pgls{dbms}. %
It is free and open source and the basis for all hands-on examples in our course.%
}%
%
In this example, you will learn several things.
You will learn about the general structure and primitives of \pglspl{rdb}.
You will learn that a \db\ consists of tables, that tables have columns and rows.
The rows represent the single records that are stored.
All the records in a table have the same attributes, each of which corresponds to a column.
You will learn how to create such tables, how to store records in them, and how to query the records from the tables.
This is the core of \pglspl{rdb}.

Talking about creating tables, storing records, and querrying them immediately leads to the question \emph{\inQuotes{How?}}
We said that we will use \postgresql.
Will the \emph{\inQuotes{How?}} that we learn be limited to this system and useless for other \pglspl{db}?
Or will it also work for \mariadb, \microsoftSqlServer, \oracleDB, \mysql, \sqlite, and so on?
Yes it will.
Because we will interact with the \db\ using the language~\sql, which basically is a programming language for \pglspl{rdb}~\cite{C20245YOQ,MS2001S1URLC,PGDG:PD:SC,SP2002STYSI2D,SPMP1998SI2TDDASVEI12T,T2018ISARD,ANSIX3135,DB2019HTTPS,ISO90751987,ISOIEC9707112023E}.
Of course, we cannot exhausively discuss \sql\ in the framework for a simple example.
But you will see some, well, examples of its use.

Finally, we will briefly touch the other topics mentioned earlier, i.e., forms, reports, and access from a programming language.
So, without further ado, here we go.
